Supplementary material to the paper 'A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2022-08-06 analysis/paper/paper.Rmd
Minor edits

The close association between Stone Age settlements in the northern parts of Scandinavia and shifting prehistoric shorelines was established by the end of the 19th century [@degeer1896], and was first applied as a dating method at the turn of the century [@brogger1905; @hollender1901]. Shoreline dating has been fundamental to Norwegian Stone Age archaeology ever since [e.g. @berg-hansen2009; @bjerck2008; @johansen1963; @bjerck1990; @breivik2014; @nummedal1923; @mikkelsen1975; @mjærum2022; @shetelig1922; @solheim2018]. The method is used both independently, and to compliment other sources of temporal data such as typological indicators or radiometric dates. However, given the coarse and fuzzy resolution of established typological frameworks, the vast amount of surveyed sites that only contain generic lithic debitage that could hail from any part of the period, and as the conditions for the preservation of organic material is typically poor in Norway, dating with reference to shoreline displacement is often the only and most precise method by which one can hope to date the sites. Shoreline dating is consequently fundamental to our understanding of the Norwegian Stone Age. This is both because it is central to the temporal framework on which our understanding of the period is based, but also because the method is only applicable so long as the societies in question have continuously settled on or close to the contemporary shoreline. Consequently, adherence or deviation from this pattern also has major implications for the socio-economic foundations of the societies in question.

>>>>

The close association between Stone Age settlements in the northern parts of Scandinavia and shifting prehistoric shorelines was established by the end of the 19th century [@degeer1896], and was first applied as a dating method at the turn of the century [@brogger1905; @hollender1901]. Shoreline dating has been fundamental to Norwegian Stone Age archaeology ever since [e.g. @berg-hansen2009; @bjerck2008; @johansen1963; @bjerck1990; @breivik2014; @olsen1984; @nummedal1923; @mikkelsen1975; @mjærum2022; @shetelig1922; @solheim2018]. The method is used both independently, and to compliment other sources of temporal data such as typological indicators or radiometric dates. However, given the coarse and fuzzy resolution of established typological frameworks, the vast amount of surveyed sites that only contain generic lithic debitage that could hail from any part of the period, and as the conditions for the preservation of organic material is typically poor in Norway, dating with reference to shoreline displacement is often the only and most precise method by which one can hope to date the sites. Shoreline dating is consequently fundamental to our understanding of the Norwegian Stone Age. This is both because it is central to the temporal framework on which our understanding of the period is based, but also because the method is only applicable so long as the societies in question have continuously settled on or close to the contemporary shoreline. Consequently, adherence or deviation from this pattern also has major implications for the socio-economic foundations of the societies in question.

### 2022-08-06 analysis/supplementary-materials/supplementary.Rmd
Minor edits

---
title: "Supplementary material to the paper 'A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast'"
author: "Isak Roalkvam"
date: "10/27/2021"
header-includes:
  - \usepackage{placeins}
output: 
  bookdown::pdf_document2:
    toc: true
    fig_caption: yes
    reference_docx: "../templates/template.docx" # Insert path for the DOCX file
    pandoc_args:
    - --lua-filter=../templates/scholarly-metadata.lua
    - --lua-filter=../templates/author-info-blocks.lua
    - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/saa.csl" # Insert path for the bib-style
nocite: | 
  @*
---

@@

The earliest ^14^C-date, from an uncertain context, is argued to be older and not match the typological indicators in the inventory from Ragnhildrød [@mjaerum2012, 50, 71]. The post-hole dated to the Late Neolithic/Bronze Age could be related to a single find of a pressure flaked object, but was not given much weight in the report [@mjaerum2012, 76].

>>>>

---
title: "Supplementary material to the paper 'A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast'"
author: 
- 'Isak Roalkvam'
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{placeins}
output: 
  bookdown::pdf_document2:
    toc: true
    fig_caption: yes
    reference_docx: "../templates/template.docx" # Insert path for the DOCX file
    pandoc_args:
    # - --lua-filter=../templates/scholarly-metadata.lua
    # - --lua-filter=../templates/author-info-blocks.lua
    - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/saa.csl" # Insert path for the bib-style
nocite: | 
  @*
---

@@

The earliest ^14^C-date, from an uncertain context, is argued to be younger and not match the typological indicators in the inventory from Ragnhildrød [@mjaerum2012, 50, 71]. The post-hole dated to the Late Neolithic/Bronze Age could be related to a single find of a pressure flaked object, but was not given much weight in the report [@mjaerum2012, 76].
